---
title: "How Does Velocity Affect Fastball Success?"
author: "Luke Stageberg"
date: "April 16, 2019"
output:
  pdf_document: default
  html_document: default
---

Naming the dataset "fb_velo" and importing it from my data folder in my Fastball_Velocity folder
```{r}
fb_velo = read.csv('../data/Fastball_Data.csv')
```


Creating a histogram to show what the average fastball velocity's were across the MLB in 2018
```{r}
hist(fb_velo$velocity, breaks = 20, main = "MLB Fastball Averages", xlab = "Pitchers' Average Velocity")
```

The first model created uses velocity as a predictor variable for batting average.
```{r}
mod_velo_ba = lm(formula = ba ~ velocity, data = fb_velo)
```

```{r}
summary(mod_velo_ba)
```

```{r}
scatter.smooth(x=fb_velo$velocity, y=fb_velo$ba, main="BA ~ Velocity", xlab="Average Velocity", ylab="Batting Average")
```

The second model created uses velocity as a predictor variable for sluggling percentage.
```{r}
mod_velo_slg = lm(formula = slg ~ velocity, data = fb_velo)
```

```{r}
summary(mod_velo_slg)
```

```{r}
scatter.smooth(x=fb_velo$velocity, y=fb_velo$slg, main="SLG ~ Velocity", xlab="Average Velocity", ylab="Slugging Percentage")
```

The third model created uses spin rate as a predictor variable for batting average
```{r}
mod_spin_ba = lm(formula = ba ~ spin_rate, data = fb_velo)
```

```{r}
summary(mod_spin_ba)
```

```{r}
scatter.smooth(x=fb_velo$spin_rate, y=fb_velo$ba, main="BA ~ Spin Rate", xlab="Average Spin Rate", ylab="Batting Average")
```

The fourth model created uses spin rate as a predictor variable for sluggling percentage.
```{r}
mod_spin_slg = lm(formula = slg ~ spin_rate, data = fb_velo)
```

```{r}
summary(mod_spin_slg)
```

```{r}
scatter.smooth(x=fb_velo$spin_rate, y=fb_velo$slg, main="SLG ~ Spin Rate", xlab="Average Spin Rate", ylab="Slugging Percentage")
```

The next step to produce another pair of models was to create a new varible that is the percentage of a swing and miss (dividing whiffs by swings). Any "na" in the data is replaced with "0"
```{r}
fb_velo$swing_and_miss_pct = with(fb_velo, whiffs / swings)
fb_velo[is.na(fb_velo)] <- 0
```

The fifth model created uses velocity as a predictor variable for swing and miss percentage.
```{r}
mod_velo_swing_and_miss =  lm(formula = swing_and_miss_pct ~ velocity, data = fb_velo)
```

```{r}
summary(mod_velo_swing_and_miss)
```

```{r}
scatter.smooth(x=fb_velo$velocity, y=fb_velo$swing_and_miss_pct, main="Swing-and-miss Percentage ~ Velocity", xlab="Average Velocity", ylab="Swing-and-miss Percentage")
```

The sixth model created uses spin rate as a predictor variable for swing and miss percentage.
```{r}
mod_spin_swing_and_miss =  lm(formula = swing_and_miss_pct ~ spin_rate, data = fb_velo)
```

```{r}
summary(mod_spin_swing_and_miss)
```

```{r}
scatter.smooth(x=fb_velo$spin_rate, y=fb_velo$swing_and_miss_pct, main="Swing-and-miss Percentage ~ Spin Rate", xlab="Average Spin Rate", ylab="Swing-and-miss Percentage")
```

The seventh model created uses velocity, spin rate, and total pitches as a predictor variable for batting average.
```{r}
mod_velo_spin_ba = lm(formula = ba ~ velocity + spin_rate + I(log(total_pitches)), data = fb_velo)
```

```{r}
summary(mod_velo_spin_ba)
termplot(mod_velo_spin_ba,partial.resid = T, se = T)
```

The eighth model created uses velocity, spin rate, and total pitches as a predictor variable for slugging percentage.
```{r}
mod_velo_spin_slg = lm(formula = slg ~ velocity + spin_rate + I(log(total_pitches)), data = fb_velo)
```

```{r}
summary(mod_velo_spin_slg)
termplot(mod_velo_spin_slg,partial.resid = T, se = T)
```


The ninth model created uses velocity, spin rate, and total pitches as a predictor variable for swing and miss percentage.
```{r}
mod_velo_spin_swing_and_miss = lm(formula = swing_and_miss_pct ~ velocity + spin_rate + I(log(total_pitches)), data = fb_velo)
```

```{r}
summary(mod_velo_spin_swing_and_miss)
termplot(mod_velo_spin_swing_and_miss,partial.resid = T, se = T)
```


